---
title: "ILC topics"
author: "Julian Trachsel"
date: "3/24/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)


library(Matrix)
library(fastTopics)
library(ggplot2)
library(cowplot)
library(Seurat)
library(tidyverse)
library(topGO)
library(biomaRt)
library(funfuns)
library(RColorBrewer)
library(kableExtra)

```


## Run the analysis at many different K values  

* See last tab for the `run_analysis()` function
```{r, eval=FALSE}

resultsK4 <- run_analysis(fit = fits$k4, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK4.rds')

resultsK5 <- run_analysis(fit = fits$k5, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK5.rds')

resultsK6 <- run_analysis(fit = fits$k6, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK6.rds')

resultsK7 <- run_analysis(fit = fits$k7, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK7.rds')

resultsK8 <- run_analysis(fit = fits$k8, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK8.rds')

resultsK9 <- run_analysis(fit = fits$k9, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK9.rds')

resultsK10 <- run_analysis(fit = fits$k10, metadata = metadata, dat=dat)
write_rds(resultsK4, path='resultsK10.rds')

```




```{r}
all_resultsK4 <- read_rds('resultsK4.rds')
all_resultsK5 <- read_rds('resultsK5.rds')
all_resultsK6 <- read_rds('resultsK6.rds')
all_resultsK7 <- read_rds('resultsK7.rds')
all_resultsK8 <- read_rds('resultsK8.rds')
all_resultsK9 <- read_rds('resultsK9.rds')
all_resultsK10 <- read_rds('resultsK10.rds')

```

## Results {.tabset}

### K=4

#### PLOTS!  

```{r, fig.width=9, fig.fullwidth=TRUE}
all_resultsK4[[1]]

```

#### Volcano plots  

```{r}
all_resultsK4[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK4[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")



```

### K=5

#### PLOTS!  

```{r, fig.width=9,fig.fullwidth=TRUE}
all_resultsK5[[1]]

```

#### Volcano plots  

```{r}
all_resultsK5[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK5[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")

```


### K=6

#### PLOTS!  

```{r, fig.width=9,fig.fullwidth=TRUE}
all_resultsK6[[1]]

```

#### Volcano plots  

```{r}
all_resultsK6[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK6[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")

```


### K=7

#### PLOTS!  

```{r,fig.width=9, fig.fullwidth=TRUE}
all_resultsK7[[1]]

```

#### Volcano plots  

```{r}
all_resultsK7[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK7[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")

```


### K=8

#### PLOTS!  

```{r,fig.width=9, fig.fullwidth=TRUE}
all_resultsK8[[1]]

```

#### Volcano plots  

```{r}
all_resultsK8[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK8[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")

```


### K=9

#### PLOTS!  

```{r, fig.width=9,fig.fullwidth=TRUE}
all_resultsK9[[1]]

```

#### Volcano plots  

```{r}
all_resultsK9[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK9[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")

```


### K=10

#### PLOTS!  

```{r, fig.width=9,fig.fullwidth=TRUE}
all_resultsK10[[1]]

```

#### Volcano plots  

```{r}
all_resultsK10[[2]]
```

#### GO-Terms for each topic  
* Top 15 GO terms for each topic
```{r}
all_resultsK10[[3]] %>%
  group_by(topic) %>% 
  slice_head(n=15) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")



```



### run_analysis code  

```{r, eval=FALSE}

# some setup first...

library(Matrix)
library(fastTopics)
library(ggplot2)
library(cowplot)
library(Seurat)
library(tidyverse)
library(topGO)
library(biomaRt)
library(funfuns)
library(RColorBrewer)



# read in data #


fits <- readRDS('./GutBlood_IntegratedILCs_K_3_K_25_topic_model_fits.rds')

dat <- readRDS('./data/GutBlood_IntegratedILCs.rds')

metadata <- 
  dat@meta.data %>% 
  mutate(ID=rownames(dat@meta.data),
         BC = sub('(.*)_([ATGC]+)-[0-9]+','\\2',barcode))


tsne_meta <- 
  tibble(ID=rownames(dat@reductions$tsne@cell.embeddings), 
         tSNE_1 = dat@reductions$tsne@cell.embeddings[,1], 
         tSNE_2 = dat@reductions$tsne@cell.embeddings[,2])

umap_meta <- 
  tibble(ID=rownames(dat@reductions$umap@cell.embeddings), 
         umap_1 = dat@reductions$umap@cell.embeddings[,1], 
         umap_2 = dat@reductions$umap@cell.embeddings[,2]) 

metadata <- 
  metadata %>%
  left_join(tsne_meta, by = 'ID') %>%
  left_join(umap_meta, by = 'ID') %>% 
  mutate(big_tissue =
           case_when(is.na(tissue)    ~  'blood', 
                     TRUE             ~   'gut'))

all(rownames(dat@meta.data) == metadata$ID)
dat@meta.data$big_tissue <- metadata$big_tissue


counts <- 
  GetAssayData(dat,assay = 'RNA', slot = 'counts') %>% 
  t() %>%   #transpose: now cells as rows, genes as columns, 
  floor() # converts to counts (round down)

counts <- counts[,colSums(counts) != 0]

# TopGO prep #
# only need to run it once to get the mapping file
# gene_IDs <- readxl::read_xlsx('UnfilteredGeneInfo.xlsx')
# 
# bm <- useMart("ensembl")
# bm <- useDataset("sscrofa_gene_ensembl", mart = bm)
# 
# EG2GO <- getBM(mart=bm, attributes=c('ensembl_gene_id', 'go_id'))
# EG2GO <- EG2GO[EG2GO$go_id != '',]
# EG2GO <- EG2GO %>% 
#   group_by(ensembl_gene_id) %>% 
#   summarise(GO=paste(go_id, sep = ' ', collapse = ',')) %>% 
#   transmute(EnsemblID=ensembl_gene_id, 
#             GO=GO)
# 
# detected_genes <- colnames(counts)
# 
# GO_gene_universe <- 
#   gene_IDs %>%
#   filter(Name %in% detected_genes) %>% 
#   left_join(EG2GO) %>% 
#   filter(!is.na(GO))
# 
# GO_gene_universe %>% 
#   dplyr::select(Name, GO) %>%
#   write_tsv('ILC_gene_to_GO.tsv')

#


# now the main function



run_analysis <- function(fit, metadata, dat){
  set.seed(1)
  
  PLOTS <- list()
  
  
  NUM_TOPICS <- ncol(fit$F)
  
  # Set topic colors:
  # use named vector so they are consistent across all figs 
  
  if (NUM_TOPICS < 10){
    topic_colors <- RColorBrewer::brewer.pal(NUM_TOPICS, 'Set1')
    names(topic_colors) <- 1:NUM_TOPICS
  } else {
    topic_colors <- colorRampPalette(colors = RColorBrewer::brewer.pal(9, 'Set1'))(NUM_TOPICS)
    names(topic_colors) <- 1:NUM_TOPICS
  }
  
  # kmeans clustering from topic based PCA (3-7 centers)
  
  pca <- prcomp(fit$L)$x
  # plot(pca)
  clusters_A <- kmeans(pca,centers = 3,iter.max = 200)$cluster
  clusters_B <- kmeans(pca,centers = 4,iter.max = 200)$cluster
  clusters_C <- kmeans(pca,centers = 5 ,iter.max = 200)$cluster
  clusters_D <- kmeans(pca,centers = 6,iter.max = 200)$cluster
  clusters_E <- kmeans(pca,centers = 7,iter.max = 200)$cluster
  
  
  kmeans_dat <- 
    tibble(
      ID=names(clusters_A), 
      kmeansA=clusters_A, 
      kmeansB=clusters_B, 
      kmeansC=clusters_C, 
      kmeansD=clusters_D, 
      kmeansE=clusters_E
    )
  metadata <- metadata %>% left_join(kmeans_dat)
  
  
  top_3_topics <- 
    fit$L %>%
    as.data.frame() %>%
    rownames_to_column(var = 'ID') %>% 
    gather(-ID, key='topic', value='proportion') %>% 
    group_by(ID) %>% 
    arrange(desc(proportion)) %>% 
    summarise(first_topic=sub('k','',topic[1]), 
              second_topic=sub('k','',topic[2]), 
              third_topic=sub('k','',topic[3]))
  
  metadata <- metadata %>% left_join(top_3_topics)
  
  # PLOT #
  PLOTS[[1]] <- 
    tibble(topic=sub('k','',names(colSums(fit$L))), 
           proportion=colSums(fit$L) / sum(colSums(fit$L))) %>% 
    ggplot(aes(x=0, y=proportion, fill=topic)) +
    geom_col(color='white') + 
    scale_fill_manual(values = topic_colors) + 
    xlim(-1,1) + 
    ggtitle('overal topic proportions') + 
    theme_cowplot()
  
  # rownames(dat@meta.data) == rownames(fits$k6$L)
  # rownames(counts) == rownames(fits$k6$L)
  
  ### Structure plots 
  
  
  # cell_type_log_lik(fit=fit, counts = counts, metadata = metadata, K=NUM_TOPICS)
  
  # PLOT
  PLOTS[[2]] <- structure_plot(fit, colors = topic_colors, topics = 1:NUM_TOPICS) + 
    ggtitle(paste('Structure plot, no groups'))
  
  # PLOT
  # using seurat clusters
  PLOTS[[3]] <- structure_plot(fit, colors = topic_colors,topics = 1:NUM_TOPICS,
                               grouping = factor(dat@meta.data$seurat_clusters),gap = 25) + 
    ggtitle('Structure plot, seurat clusters')
  
  # PLOT
  # using clusters generated from topics
  # HOW TO DETERMINE OPTIMAL CLUSTS?
  PLOTS[[4]] <- structure_plot(fit,colors = topic_colors,
                               grouping = factor(clusters_A),gap = 25, topics = 1:NUM_TOPICS,) + 
    ggtitle('Kmeans clusters based on topics, num_clusts = 3')
  
  
  # PLOT
  PLOTS[[5]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=factor(kmeansA))) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    # scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by kmeans clusters') + 
    facet_wrap(~big_tissue)
  
  
  # PLOT
  PLOTS[[6]] <- structure_plot(fit,colors = topic_colors,
                               grouping = factor(clusters_B),gap = 25, topics = 1:NUM_TOPICS,) + 
    ggtitle(paste('Kmeans clusters based on topics, num_clusts = 4'))
  
  # PLOT
  PLOTS[[7]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=factor(kmeansB))) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    # scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by kmeans clusters') + 
    facet_wrap(~big_tissue)
  
  
  # PLOT
  PLOTS[[8]] <- structure_plot(fit,colors = topic_colors,
                               grouping = factor(clusters_C),gap = 25, topics = 1:NUM_TOPICS,) + 
    ggtitle(paste('Kmeans clusters based on topics, num_clusts = 5'))
  
  
  # PLOT
  PLOTS[[9]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=factor(kmeansC))) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    # scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by kmeans clusters') + 
    facet_wrap(~big_tissue)
  
  
  # PLOT
  PLOTS[[10]] <- structure_plot(fit,colors = topic_colors,
                                grouping = factor(clusters_D),gap = 25, topics = 1:NUM_TOPICS,) + 
    ggtitle(paste('Kmeans clusters based on topics, num_clusts = 6 '))
  
  
  # PLOT
  PLOTS[[11]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=factor(kmeansD))) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    # scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by kmeans clusters') + 
    facet_wrap(~big_tissue)
  
  
  # PLOT
  PLOTS[[12]] <- structure_plot(fit,colors = topic_colors,
                                grouping = factor(clusters_E),gap = 25, topics = 1:NUM_TOPICS,) + 
    ggtitle(paste('Kmeans clusters based on topics, num_clusts = 7 '))
  
  
  # PLOT
  PLOTS[[13]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=factor(kmeansE))) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    # scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by kmeans clusters') + 
    facet_wrap(~big_tissue)
  
  
  
  # PLOT
  # structure plot using sample type groupings
  PLOTS[[14]] <- structure_plot(fit, colors = topic_colors, 
                                grouping = factor(metadata$big_tissue), gap = 25) + 
    ggtitle('topics by tissue')
  
  
  ### TSNE colored by most abundant topics ###
  # browser()
  # PLOT
  PLOTS[[15]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=first_topic)) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by max topic') + 
    facet_wrap(~big_tissue)
  # PLOT
  PLOTS[[16]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=second_topic)) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by 2nd topic') + 
    facet_wrap(~big_tissue)
  # PLOT
  PLOTS[[17]] <- metadata %>%
    ggplot(aes(x=tSNE_1, y=tSNE_2, color=third_topic)) + 
    geom_point(alpha=.5) + 
    theme_cowplot() + 
    scale_color_manual(values=topic_colors) + 
    ggtitle('tSNE, each cell colored by 3nd topic') + 
    facet_wrap(~big_tissue)
  
  
  
  
  ######### Diffabund?
  
  dfa_out <- diff_count_analysis(fit,counts)
  
  all_topics_volcano_plots <- function(dfa_out){
    res <- list()
    for (K in 1:NUM_TOPICS){
      p <- volcano_plot(diff_count_result = dfa_out, k=K)
      res[[K]] <- p
    }
    return(res)
  }
  
  VOLCs <- all_topics_volcano_plots(dfa_out)
  
  # PLOTS
  VOLCs
  
  
  Topic_enriched_genes <- 
    dfa_out$beta %>%
    as.data.frame() %>% 
    rownames_to_column(var = 'Gene') %>% 
    gather(-Gene, key='topic', value='LFC') %>% 
    filter(LFC > 0) %>% 
    group_by(topic) %>%
    nest() %>% 
    mutate(enrich_val = map_dbl(data, ~ quantile(.x$LFC, .99)), 
           enrich_genes= map2(.x=data, .y=enrich_val, ~ filter(.x, LFC > .y))) 
  
  
  
  
  GO_terms_4_topics <- 
    Topic_enriched_genes %>% 
    dplyr::select(topic, enrich_genes) %>% 
    mutate(topGO_res=map(.x = enrich_genes, ~ topGO_wrapper(myInterestingGenes = .x$Gene, mapping_file = 'ILC_gene_to_GO.tsv', ont = 'BP')), 
           filtered_res=map(.x = topGO_res, ~ filter(.x, pval < 0.05))) %>% 
    dplyr::select(topic, filtered_res)
  
  
  sig_GO_terms_4_topics <- GO_terms_4_topics %>% unnest(filtered_res)
  
  
  all_results <- list(PLOTS, VOLCs, sig_GO_terms_4_topics)
  return(all_results)
}





```


