---
title: "Selected_topics"
author: "Julian Trachsel"
date: "5/19/2021"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.full.width = TRUE)
knitr::opts_chunk$set(fig.dim=c(8,6))


library(tidyverse)
library(fastTopics) 
library(Seurat)
library(scales)
library(funfuns)
library(knitr)
library(kableExtra)

source('./scripts/analyze_fits.R')
source('./scripts/selected_fits.R')


```



CD4 only: topics k = 3, spatial DGE modules = 3  
gd/CD8 only: topics = 3, spatial DGE modules = 4 and modules = 6  
ILC only: topics = 3, spatial DGE modules = 3  
B only: topics = 6, spatial DGE modules = 9  
Myeloid only: topics = 5, spatial = don't know yet; skip for now  
Non-immune only: topics = 6, spatial DGE modules = 9  

## Different Subsets of Interest {.tabset .tabset-fade .tabset-pills}  

  
### CD4: topics k = 3 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r, fig.fullwidth=TRUE}
results$CD4$feature_plot
results$CD4$violin_plot
results$CD4$struct_plot_1

```


#### Volcano Plots
```{r}
results$CD4$volcano_plots
```


#### Topic Enriched Genes
```{r}

results$CD4$topics_sig_genes %>% filter(LFC >0)%>% arrange(desc(Zscore)) %>%
  kable() %>%
  kable_styling(bootstrap_options = 'striped')
```


#### GO Terms 

```{r}
results$CD4$GO_terms %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```




  
### gdCD8: topics k = 3 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r, fig.fullwidth=TRUE}
results$gdCD8$feature_plot
results$gdCD8$violin_plot
results$gdCD8$struct_plot_1

```


#### Volcano Plots
```{r}
results$gdCD8$volcano_plots
```


#### Topic Enriched Genes
```{r}

results$gdCD8$topics_sig_genes %>% filter(LFC >0)%>% arrange(desc(Zscore)) %>%
  kable() %>%
  kable_styling(bootstrap_options = 'striped')
```


#### GO Terms 

```{r}
results$gdCD8$GO_terms %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```


### ILC: topics k = 3 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r, fig.fullwidth=TRUE}
results$ILC$feature_plot
results$ILC$violin_plot
results$ILC$struct_plot_1

```


#### Volcano Plots
```{r}
results$ILC$volcano_plots
```


#### Topic Enriched Genes
```{r}

results$ILC$topics_sig_genes %>% filter(LFC >0)%>% arrange(desc(Zscore)) %>%
  kable() %>%
  kable_styling(bootstrap_options = 'striped')
```


#### GO Terms 

```{r}
results$ILC$GO_terms %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```


### BCell: topics k = 6 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r, fig.fullwidth=TRUE}
results$BCell$feature_plot
results$BCell$violin_plot
results$BCell$struct_plot_1

```


#### Volcano Plots
```{r}
results$BCell$volcano_plots
```


#### Topic Enriched Genes
```{r}

results$BCell$topics_sig_genes %>% filter(LFC >0)%>% arrange(desc(Zscore)) %>%
  kable() %>%
  kable_styling(bootstrap_options = 'striped')
```


#### GO Terms 

```{r}
results$BCell$GO_terms %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```



  
### Myeloid: topics k = 5 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r, fig.fullwidth=TRUE}
results$Myeloid$feature_plot
results$Myeloid$violin_plot
results$Myeloid$struct_plot_1

```


#### Volcano Plots
```{r}
results$Myeloid$volcano_plots
```


#### Topic Enriched Genes
```{r}

results$Myeloid$topics_sig_genes %>% filter(LFC >0)%>% arrange(desc(Zscore)) %>%
  kable() %>%
  kable_styling(bootstrap_options = 'striped')
```


#### GO Terms 

```{r}
results$Myeloid$GO_terms %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```



### NonImmune: topics k = 6 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r, fig.fullwidth=TRUE}
results$NonImmune$feature_plot
results$NonImmune$violin_plot
results$NonImmune$struct_plot_1

```


#### Volcano Plots
```{r}
results$NonImmune$volcano_plots
```


#### Topic Enriched Genes
```{r}

results$NonImmune$topics_sig_genes %>% filter(LFC >0)%>% arrange(desc(Zscore)) %>%
  kable() %>%
  kable_styling(bootstrap_options = 'striped')
```


#### GO Terms 

```{r}
results$NonImmune$GO_terms %>% kable() %>% kable_styling(bootstrap_options = 'striped')
```

