---
title: "Gut Blood ILCs - Topic Model Exploration"
author: "Julian Trachsel"
date: "5/5/2021"
output: html_document
---
```{r, warning=F, message=F, results='hide'}
library(tidyverse)
library(fastTopics) 
library(Seurat)
library(scales)

source('./scripts/analyze_fits.R')

SEURAT <- read_rds('./data/GutBlood_IntegratedILCs.rds')
FITS <- read_rds('./topic_model_fits/GutBlood_IntegratedILCs_K_3_K_25_topic_model_fits.rds')

GROUP <- 'GutBlood_ILC' # probably can find a way to cut this

results <- analyze_fits(SEURAT = SEURAT, FITS = FITS[1:8], GROUP = GROUP, SEED=2)

```


## tabs {.tabset .tabset-fade .tabset-pills}  

Different levels of topic model K below:  

### K = 3 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k3$result$feature_plot
results$k3$result$violin_plot
results$k3$result$struct_plot_1
results$k3$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k3$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k3$result$KMEANS_RES$KM4
```
##### KMeans K=5  

```{r}
results$k3$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k3$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k3$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k3$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k3$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k3$result$KMEANS_RES$KM10
```

### K = 4 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k4$result$feature_plot
results$k4$result$violin_plot
results$k4$result$struct_plot_1
results$k4$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k4$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k4$result$KMEANS_RES$KM4
```
##### KMeans K=5  

```{r}
results$k4$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k4$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k4$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k4$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k4$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k4$result$KMEANS_RES$KM10
```





### K = 5 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k5$result$feature_plot
results$k5$result$violin_plot
results$k5$result$struct_plot_1
results$k5$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k5$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k5$result$KMEANS_RES$KM4
```
##### KMeans K=5  

```{r}
results$k5$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k5$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k5$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k5$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k5$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k5$result$KMEANS_RES$KM10
```





### K = 6 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k6$result$feature_plot
results$k6$result$violin_plot
results$k6$result$struct_plot_1
results$k6$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k6$result$KMEANS_RES$KM3
```
  
##### KMeans K=4  
```{r}
results$k6$result$KMEANS_RES$KM4
```
  
##### KMeans K=5  
```{r}
results$k6$result$KMEANS_RES$KM5
```
  
##### KMeans K=6  
```{r}
results$k6$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k6$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k6$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k6$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k6$result$KMEANS_RES$KM10
```





### K = 7 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k7$result$feature_plot
results$k7$result$violin_plot
results$k7$result$struct_plot_1
results$k7$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k7$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k7$result$KMEANS_RES$KM4
```

##### KMeans K=5  
```{r}
results$k7$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k7$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k7$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k7$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k7$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k7$result$KMEANS_RES$KM10
```





### K = 8 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k8$result$feature_plot
results$k8$result$violin_plot
results$k8$result$struct_plot_1
results$k8$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k8$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k8$result$KMEANS_RES$KM4
```

##### KMeans K=5  
```{r}
results$k8$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k8$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k8$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k8$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k8$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k8$result$KMEANS_RES$KM10
```





### K = 9 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k9$result$feature_plot
results$k9$result$violin_plot
results$k9$result$struct_plot_1
results$k9$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k9$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k9$result$KMEANS_RES$KM4
```

##### KMeans K=5  
```{r}
results$k9$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k9$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k9$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k9$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k9$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k9$result$KMEANS_RES$KM10
```





### K = 10 {.tabset .tabset-fade .tabset-pills}  

#### General Plots  
```{r}
results$k10$result$feature_plot
results$k10$result$violin_plot
results$k10$result$struct_plot_1
results$k10$result$struct_plot_2
```


#### KMEANS {.tabset .tabset-fade .tabset-pills}  

##### KMeans K=3  
```{r}
results$k10$result$KMEANS_RES$KM3
```

##### KMeans K=4  
```{r}
results$k10$result$KMEANS_RES$KM4
```

##### KMeans K=5  
```{r}
results$k10$result$KMEANS_RES$KM5
```

##### KMeans K=6  
```{r}
results$k10$result$KMEANS_RES$KM6
```   

##### KMeans K=7  
```{r}
results$k10$result$KMEANS_RES$KM7
```

##### KMeans K=8  
```{r}
results$k10$result$KMEANS_RES$KM8
```

##### KMeans K=9  
```{r}
results$k10$result$KMEANS_RES$KM9
```

##### KMeans K=10  
```{r}
results$k10$result$KMEANS_RES$KM10
```

```{r}
# 
# 
# for (KVAL in names(results)){
# 
#   
# 
#   i <- which(metadata$result_genome == result_genome)
# 
#   exper <- metadata$experiment[i]
# 
#   don <- metadata$donor_genome[i]
# 
#   rec <- metadata$recipient_genome[i]
# 
#   resu <- metadata$result_genome[i]
# 
#   exp_name <- paste(exper,':', don, '+', rec, '=', resu )
# 
# 
# 
#   cat("  \n###",  exp_name, "\n")
# 
# 
# 
#   RESULTS[[result_genome]] %>%
# 
#   select(c(1,6, 12, 16, 19, 20, 26)) %>%
# 
#   kable(caption = 'Transferred genomic islands') %>%
# 
#   kable_styling(bootstrap_options = 'striped') %>% print()
# 
# 
# 
# 
# 
#   plot_islands(RESULTS[[result_genome]], clust_level = quat_cluster) %>% print()
# 
# 
# 
#   TRANSFERRED[[result_genome]] %>%
# 
#   select(Gene, all_Qclusters, Annotation, 22:24) %>%
# 
#   kable(caption = 'Transferred Genes') %>%
# 
#   kable_styling(bootstrap_options = 'striped') %>% print()
# 
# 
# 
#   DELETED[[result_genome]] %>%
# 
#   select(Gene, all_Qclusters, Annotation, 22:24) %>%
# 
#   kable(caption = 'Deleted Genes') %>%
# 
#   kable_styling(bootstrap_options = 'striped') %>% print()
# 
#   cat(" END", exp_name, "  \n")
# 
# 
# 
#    cat("  \n*****
# 
# 
# 
# <br/>
# 
# <br/>
# 
# <br/>
# 
#   \n")
# 
# }
# 

```