---
title: "Reference-based Label Prediction and Mapping of Porcine Ileal Cells: Creation of Porcine Ileum Query Datasets"
author: "Jayne Wiarda"
date: "5Apr2021"
output: 
  github_document: 
    keep_html: TRUE
---

Since there are already some pretty spectacular and well-annotated scRNA-seq datasets available in literature, we can utilize these external datasets as references we can compare to our porcine ileum scRNA-seq data, which we treat as a dataset to query. We can perform reference-based label transfer to predict the most similar cellular counterparts in already annotated datasets we use as references (Which reference cells are our query cells most similar to?), and we can further calculate mapping scores to see how well query data (our porcine ileum scRNA-seq data) is represented by the reference dataset (How well does the reference data truly resemble the query data?).

Unfortunately, we are the first to create a scRNA-seq dataset of cells in the porcine intestine, so we have to compare to other reference datasets either within the same species but across anatomical locations or within the same tissue but across species. To compare across species, we must also adjust our gene count matrices to include only gene orthologs in our query data. Thus, we create three 'versions' of our query datasets:

* Query data with all expressed genes and porcine gene names, as used in previous analyses of the porcine ileum scRNA-seq data, referred to as query_SsIleum_Wiarda
* Query data with only one-to-one human-to-pig gene orthologs and genes converted to human gene names, referred to as query_SsIleum_Wiarda_humanized
* Query data with only one-to-one mouse-to-pig gene orthologs and genes converted to murine gene names, referred to as query_SsIleum_Wiarda_murinized

Our query_SsIleum_Wiarda_humanized data can then be directly compared to human reference datasets that have also been filtered to contain only one-to-one human-to-pig gene orthologs (with human gene names present in the reference), while our query_SsIleum_Wiarda_murinized data can be compared directly to murine reference datasets filtered to contain only one-to-one mouse-to-pig gene orthologs (with murine gene names present in the reference).

In this analysis, we create query datasets that can be compared directly to other porcine scRNA-seq reference datasets (query_SsIleum_Wiarda).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

### Load required software packages

The R package, Seurat (Stuart & Butler et al. 2019; Hao et al. 2020), is utilized to perform label transfer and mapping. Note, this analysis is done with an upgraded version of Seurat compared to other analyses, and is also performed on another machine. We also utilize the R package, future (Bengtsson 2020), to parallelize some of the steps performed herein. Refer to sessionInfo() at the bottom of the page for the R and package versions used and additional information.

Additional packages used: ggplot2 (Wickham 2016)

```{r, message = FALSE}
library(ggplot2)
library(Seurat)
library(future)
```

## Create porcine ileum query datasets

### Load relevant data:

Load in the scRNA-seq porcine ileum Seurat object we've been working with in previous analyses:

```{r, warning = FALSE, error = FALSE}
SsIleum_Wiarda <- readRDS("/Users/jayne.wiarda/Desktop/scRNAseq/scRNAseqIleumAtlas/Ileum/Seurat/IleumAtlasAll.rds") # read in our pig ileum atlas dataset
```

### Allow parallelization 

```{r, warning = FALSE, error = FALSE}
plan("multiprocess", workers = 4) # change plan to allow parallelization with 4 workers
options(future.globals.maxSize = 5000 * 1024^2)
```

### Remove unneccessary information

Remove SCT and integrated assays, and remove all dimensionality reductions of the data:

```{r, warning = FALSE, error = FALSE}
SsIleum_Wiarda <- DietSeurat(SsIleum_Wiarda, 
                                   assays = 'RNA', # remove SCT and integrated assays
                                   dimreducs = NULL) # remove former dimensionality reductions
```

### Normalization, scaling, and dimensionality reduction

Split the data into individual samples:

```{r, warning = FALSE, error = FALSE, message = FALSE}
SsIleum_Wiarda.list <- SplitObject(SsIleum_Wiarda, 
                                   split.by = "SampleID") # split into the original samples that were processed for scRNA-seq
```

At this point, we treat data from each sample as an independent entity, giving us six query datasets that need to be processed. For each sample, normalize the data, find the top 2,000 most variable genes, scale the data, and run PCA for the first 50 PCs:

```{r, warning = FALSE, error = FALSE, message = FALSE}
for (i in 1:length(SsIleum_Wiarda.list)) { # for each sample individually, let's normalize the data and find the 2000 most highly variable features, scale the data, and find top 50 PCs
  SsIleum_Wiarda.list[[i]] <- NormalizeData(SsIleum_Wiarda.list[[i]], 
                                            verbose = TRUE, 
                                            normalization.method = "LogNormalize", 
                                            scale.factor = 10000, 
                                            assay = "RNA")
  SsIleum_Wiarda.list[[i]] <- FindVariableFeatures(SsIleum_Wiarda.list[[i]], 
                                                   selection.method = "vst", 
                                                   nfeatures = 2000, 
                                                   verbose = TRUE)
  SsIleum_Wiarda.list[[i]] <- ScaleData(SsIleum_Wiarda.list[[i]]) # scale the data
  SsIleum_Wiarda.list[[i]] <- RunPCA(SsIleum_Wiarda.list[[i]], # calculate 100 PCs
                                     npcs = 50, 
                                     verbose = TRUE)
}
```

### Save the query datasets as a Seurat object:

```{r, warning = FALSE, error = FALSE}
saveRDS(SsIleum_Wiarda.list, '/Users/jayne.wiarda/Desktop/scRNAseq/scRNAseqIleumAtlas/Ileum/QueryData/query_SsIleum_Wiarda.rds')
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References

Integrated analysis of multimodal single-cell data. Yuhan Hao, Stephanie Hao, Erica Andersen-Nissen, William M. Mauck III, Shiwei Zheng, Andrew Butler, Maddie J. Lee, Aaron J. Wilk, Charlotte Darby, Michael Zagar, Paul Hoffman, Marlon Stoeckius, Efthymia Papalexi, Eleni P. Mimitou, Jaison Jain, Avi Srivastava, Tim Stuart, Lamar B. Fleming, Bertrand Yeung, Angela J. Rogers, Juliana M. McElrath, Catherine A. Blish, Raphael Gottardo, Peter Smibert, Rahul Satija. bioRxiv 2020.10.12.335331; doi: https://doi.org/10.1101/2020.10.12.335331

H. Bengtsson, A Unifying Framework for Parallel and Distributed Processing in R using Futures, arXiv:2008.00553, 2020
 
Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
