---
title: "Characterization of Ileal T Cells & Innate Lymphoid Cells: Overall Interpretations & Cell Calling"
author: "Jayne Wiarda"
date: "1Apr2021"
output: 
  github_document: 
    keep_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created a data subset consisting of only T/ILC lineage lymphocytes and performed further analyses to characterize the data. Let's now visualize the data a few other ways and finally decide on more highly-resolved cell identites.

Sample nomenclature is as follows:

* IPP = ileum with only Peyer's patch regions present
* NoPP = ileum with only regions without Peyer's patches present
* X2I = ileum with both regions with and without Peyer's patches present; corresponding to a whole ileum cross section
* Pig 1 = IPP1, NoPP1, X2I1
* Pig 2 = IPP2, NoPP2, X2I2

In this analysis, we end up classifying our T/ILC lineage lymphocytes into groupings as follows:

* CD4 ab T cells = clusters 12, 15, 26, 41, 46
* Mixed CD4 & CD8 ab T cells (naive) = cluster 24
* CD8 ab T cells = clusters 3, 7, 14, 17, 37
* Mixed gd & CD8 ab T cells = clusters 19, 34, 35, 47
* gd T cells = clusters 6, 23, 29, 31, 32, 51
* ILCs = clusters 1, 18, 43, 44, 53

## Overall Interpretations & Cell Calling

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

```{r, message = FALSE}
library(Seurat)         
```

### Import relevant data 

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
ilT <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_TILConly.rds") 
DefaultAssay(ilT) <- "RNA" # set our default assay to RNA for subsequent work
Idents(ilT) <- ilT$seurat_clusters # change our default IDs to our cluster assignments
```

Also bring in the number of 'significant' PCs that we calculated for this dataset previously:

```{r, warning = FALSE, error = FALSE}
PCdims <- 1:10
```

### Not done, but include all the following:

# Plot some canonical T cell genes:
FeaturePlot(ilT,
            features = c('CD4', 'CD8B', 'TRDC', 'CD3E', 'CD2', 'CD8A'),
            reduction = 'tsne',
            cols = c("grey90", "purple4"),
            ncol = 3) & NoLegend() & NoAxes()

DotPlot(ilT, 
        features = c('CD4', 'CD8B', 'TRDC', 'CD3E', 'CD2', 'PTPRC', 'CD79A', 'ENSSSCG00000028461'), 
        cols = c("gold", "red"), 
        col.min = -1, 
        col.max = 2,
        dot.scale = 6,
        assay = "RNA") +
  RotatedAxis()

# Reorder cells:
Idents(ilT) <- ilT$seurat_clusters
levels(ilT) <- c('12', '15', '26', '41', '46', 
                 '24', 
                 '3', '7', '14', '17', '37',
                 '19', '34', '35', '47',
                 '6', '23', '29', '31', '32', '51',
                 '1', '18', '43', '44', '53') # Reorder the clusters based on putative cell type IDs we came up with from looking at the data
ilT$neworder <- Idents(ilT)
Idents(ilT) <- ilT$neworder

DotPlot(ilT, 
        features = c('CD4', 'CD8B', 'TRDC', 'CD3E', 'CD2', 'CD8A', 'PTPRC', 'CD79A', 'ENSSSCG00000028461'), 
        cols = c("gold", "red"), 
        col.min = -1, 
        col.max = 2,
        dot.scale = 6,
        assay = "RNA",
        group.by = 'neworder') +
  RotatedAxis()

Idents(ilT) <- ilT$neworder
TILCtypes <- c(rep('CD4 ab T cell', 5), 'CD4/CD8 mix', rep('CD8 ab T cell', 5),
  rep('CD8/gd mix', 4), rep('gd T cell', 6), 
  rep('ILC', 5)) # Rename clusters based on phenotype IDs
ilT$TILCtypes <- ilT$neworder
Idents(ilT) <- ilT$TILCtypes
names(TILCtypes) <- levels(ilT) # assign TILCtypes to cluster numbers
ilT <- RenameIdents(ilT, TILCtypes) # change dataset identity to cell types in new Seurat object
ilT$TILCtypes <- Idents(ilT)

DimPlot(ilT,
        reduction = 'tsne',
        group.by = 'TILCtypes', 
        cols = c('violetred', 'darkorange', 'gold', 'limegreen', 'dodgerblue3', 'purple4'))

### Assign cell identities:

```{r, warning = FALSE, error = FALSE, eval = TRUE}
saveRDS(ilT, "/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_TILConly.rds")
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References

Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).
