---
title: "Characterization of Ileal ILCs: Hierarchical Clustering"
author: "Jayne Wiarda"
date: "9Apr2021"
output: 
  github_document: 
    keep_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created a data subset consisting of only ILCs. Let's now perform hierarchical clustering analysis on just these clusters of cells.

Sample nomenclature is as follows:

* IPP = ileum with only Peyer's patch regions present
* NoPP = ileum with only regions without Peyer's patches present
* X2I = ileum with both regions with and without Peyer's patches present; corresponding to a whole ileum cross section
* Pig 1 = IPP1, NoPP1, X2I1
* Pig 2 = IPP2, NoPP2, X2I2

## Hierarchical clustering of cell clusters

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

```{r, message = FALSE}
library(Seurat)         
```

### Import relevant data 

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
ilILC <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_ILConly.rds") 
DefaultAssay(ilILC) <- "RNA" # set our default assay to RNA for subsequent work
Idents(ilILC) <- ilILC$seurat_clusters # change our default IDs to our cluster assignments
```

Also bring in the number of 'significant' PCs that we calculated for this dataset previously:

```{r, warning = FALSE, error = FALSE}
PCdims <- 1:4
```

### Calculate hierarchical clustering

Build a phylogenetic tree, utilizing seurat_clusters as our groups and using the pre-determined significant number of PCs to specify the dimensionality of our data to use:

```{r, warning = FALSE, error = FALSE}
ilILC <- BuildClusterTree(ilILC, 
                       dims = PCdims, 
                       assay = "PCA")
```

Now let's look at our cluster tree:

```{r, warning = FALSE, error = FALSE, fig.width = 20}
PlotClusterTree(ilILC, 
                edge.width = 3) # plot tree with node labels
```

Let's remove node labels and plot the tree again:

```{r, warning = FALSE, error = FALSE, fig.width = 20}
data.tree <- Tool(object = ilILC, 
                  slot = "BuildClusterTree") 
ape::plot.phylo(x = data.tree, 
                direction = "downwards", # plot the tree without node labels
                edge.width = 1.5)
```

Let's reorder some of the branches while still maintaining our heirarchy:

```{r, warning = FALSE, error = FALSE, fig.width = 20}
data.tree <- ape::rotateConstr(data.tree, c('43', '44', '1', '18', '53'))
plot(data.tree, direction = 'downwards', edge.width = 1.5, font = 1)
```

For future analyses, we may want to order our clusters by this phylogenetic order, so let's store that information in a metadata slot for our Seurat object, called 'phyloorder':

```{r, warning = FALSE, error = FALSE}
levels(ilILC) <- rev(c('43', '44', '1', '18', '53')) 
ilILC$phyloorder <- Idents(ilILC)
```

### Save our data

Since we've added some new meta data information to our dataset, let's save our Seurat object as an .rds again:

```{r, warning = FALSE, error = FALSE, eval = TRUE}
saveRDS(ilILC, "/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_ILConly.rds")
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References

Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).
