---
title: "Characterization of Ileal Myeloid Lineage Leukocytes: Visualizing Prediction & Mapping Results"
author: "Jayne Wiarda"
date: "15Apr2021"
output: 
  github_document: 
    keep_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created a data subset consisting of only myeloid lineage leukocytes. Let's now overlay preference-based prediction & mapping information back onto our myeloid data.

Sample nomenclature is as follows:

* IPP = ileum with only Peyer's patch regions present
* NoPP = ileum with only regions without Peyer's patches present
* X2I = ileum with both regions with and without Peyer's patches present; corresponding to a whole ileum cross section
* Pig 1 = IPP1, NoPP1, X2I1
* Pig 2 = IPP2, NoPP2, X2I2

## Visualizing reference-based cell predictions and mapping

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

Other required packages: ggplot2 (Wickham 2016), scales (Wickham & Seidel 2020)

```{r, message = FALSE}
library(Seurat)  
library(scales)
library(ggplot2)
```

### Import relevant data 

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
ilM <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_MyeloidOnly.rds") 
DefaultAssay(ilM) <- "RNA"
```

### Visualize mapping scores

Plot mapping scores overlaid onto our t-SNE plot:

```{r, warning = FALSE, error = FALSE, message = FALSE, fig.width = 18}
FeaturePlot(ilM, 
            features = c('SsPBMC_MappingScores', 
                         'MmIleum_MappingScores', 
                         'HsIleum_MappingScores'),
            reduction = 'tsne', # change to 'umap' to overlay onto UMAP plot
            ncol = 3) & 
  scale_color_gradientn( colours = c('yellow', 'orange', 'red', 'red4'),  limits = c(0.5, 1), oob = squish) & 
  NoAxes() & NoLegend()
```

### Visualize selected cell type prediction scores

We selected and plotted overlaid prediction scores for myeloid populations found in reference data.

From murine ileum reference cell types:

```{r, warning = FALSE, error = FALSE, fig.width = 15, message = FALSE, fig.height = 11}
FeaturePlot(ilM, 
            features = c('MmIleum_CellType_prediction.score.pDC',
                         'MmIleum_CellType_prediction.score.DC..CD103posCD11bpos.',
                         'MmIleum_CellType_prediction.score.DC..CD103posCD11bneg.',
                         'MmIleum_CellType_prediction.score.DC..CD103neg.C1.',
                         'MmIleum_CellType_prediction.score.DC..CD103neg.C2.',
                         'MmIleum_CellType_prediction.score.Macrophage',
                         'MmIleum_CellType_prediction.score.Mast.cell',
                         'MmIleum_CellType_prediction.score.Neutrophil',
                         'MmIleum_CellType_prediction.score.Basophil'),
            reduction = 'tsne', # change to 'umap' to overlay onto UMAP plot
            ncol = 4) & 
  scale_color_gradientn( colours = c('grey90', 'blue3'),  limits = c(0, 0.5), oob = squish) & 
  NoAxes() & NoLegend()
```

From human ileum reference cell types:

```{r, warning = FALSE, error = FALSE, fig.width = 15, message = FALSE, fig.height = 7}
FeaturePlot(ilM, 
            features = c('HsIleum_CellType_prediction.score.pDC',
                         'HsIleum_CellType_prediction.score.cDC1',
                         'HsIleum_CellType_prediction.score.cDC2',
                         'HsIleum_CellType_prediction.score.activated.DC',
                         'HsIleum_CellType_prediction.score.Cycling.myeloid.cells',
                         'HsIleum_CellType_prediction.score.Macrophage',
                         'HsIleum_CellType_prediction.score.Monocyte',
                         'HsIleum_CellType_prediction.score.activated.mast.cells'),
            reduction = 'tsne', # change to 'umap' to overlay onto UMAP plot
            ncol = 3) & 
  scale_color_gradientn( colours = c('grey90', 'blue3'),  limits = c(0, 0.5), oob = squish) & 
  NoAxes() & NoLegend()
```

From porcine PBMC reference cell types:

```{r, warning = FALSE, error = FALSE, fig.width = 15, message = FALSE, fig.height = 7}
FeaturePlot(ilM, 
            features = c('SsPBMC_CellType_prediction.score.pDC',
                         'SsPBMC_CellType_prediction.score.cDC',
                         'SsPBMC_CellType_prediction.score.Monocytes'),
            reduction = 'tsne', # change to 'umap' to overlay onto UMAP plot
            ncol = 3) & 
  scale_color_gradientn( colours = c('grey90', 'blue3'),  limits = c(0, 0.5), oob = squish) & 
  NoAxes() & NoLegend()
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References

Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
 
Hadley Wickham and Dana Seidel (2020). scales: Scale Functions for Visualization. R package version 1.1.1. https://CRAN.R-project.org/package=scales
  