---
title: "Adding cell annotations to T/ILC dataset"
author: "Jayne Wiarda"
date: "02June2021"
output: 
  github_document: 
    keep_html: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created our final annotations for our T/ILC data and need to add these and visualize.

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019), ggplot2 (Wickham 2016), scales (Wickham & Seidel 2020), dplyr (Wickham et al. 2020), readxl (Wickham & Bryan 2019), reshape2 (Wickham 2007). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

```{r, message = FALSE}
library(Seurat)
library(ggplot2)
library(scales)
library(dplyr)
library(readxl)
library(reshape2)
```

### Import relevant data

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
ilT <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_TILConly.rds") 
DefaultAssay(ilT) <- "RNA" # set our default assay to RNA for subsequent work
Idents(ilT) <- ilT$seurat_clusters # change our default IDs to our cluster assignments
```

Read in cell annotation files:

```{r, warning = FALSE, error = FALSE}
CD4ann <- read.table('/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Annotations/CD4T.txt')
gdCD8ann <- read.table('/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Annotations/gdCD8T.txt')
ILCann <- read.table('/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Annotations/ILC.txt')
```

## Add cell annotations to the dataset

```{r, warning = FALSE, error = FALSE}
bcs <- as.data.frame(colnames(ilT))
colnames(bcs) <- 'barcode'
bcs$cellID <- rep('_', nrow(bcs))
df <- as.data.frame(ilT$seurat_clusters)
# Cluster-based cell annotations not provided in the annotation .txt files:
naiveT <- rownames(subset(df, ilT$seurat_clusters == '24'))

ActivatedCD4T <- subset(CD4ann, cellID == 'Activated CD4 ab T')
CyclingCD4T <- subset(CD4ann, cellID == 'Cycling CD4 ab T')
FollicularCD4T <- subset(CD4ann, cellID == 'Follicular CD4 ab T')
CD8_activated <- subset(gdCD8ann, cellID == 'Activated CD8 ab T')
CD8_cycling <- subset(gdCD8ann, cellID == 'Cycling CD8 ab T')
CD8_cytotoxic <- subset(gdCD8ann, cellID == 'Cytotoxic CD8 ab T')
gd_activated <- subset(gdCD8ann, cellID == 'Activated gd T')
gd_CD2neg <- subset(gdCD8ann, cellID == 'CD2neg gd T')
gd_cytotoxic <- subset(gdCD8ann, cellID == 'Cytotoxic gd T')
gd_cycling <- subset(gdCD8ann, cellID == 'Cycling gd T')
gd_SELLpos <- subset(gdCD8ann, cellID == 'SELLhi gd T')
ILC1 <- subset(ILCann, cellID == 'Activated group 1 ILCs')
ILC3 <- subset(ILCann, cellID == 'Group 3 ILCs')
ILC1_cytotoxic <- subset(ILCann, cellID == 'Cytotoxic group 1 ILCs')
ILC1_cycling <- subset(ILCann, cellID == 'Cycling group 1 ILCs')

bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% naiveT, 'Naive CD4/CD8 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% ActivatedCD4T$barcode, 'Activated CD4 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% CyclingCD4T$barcode, 'Cycling CD4 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% FollicularCD4T$barcode, 'Follicular CD4 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% CD8_activated$barcode, 'Activated CD8 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% CD8_cycling$barcode, 'Cycling CD8 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% CD8_cytotoxic$barcode, 'Cytotoxic CD8 ab T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% gd_activated$barcode, 'Activated gd T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% gd_CD2neg$barcode, 'CD2neg GD T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% gd_cytotoxic$barcode, 'Cytotoxic gd T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% gd_cycling$barcode, 'Cycling gd T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% gd_SELLpos$barcode, 'SELLhi gd T cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% ILC1$barcode, 'Activated group 1 ILCs'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% ILC3$barcode, 'Group 3 ILCs'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% ILC1_cytotoxic$barcode, 'Cytotoxic group 1 ILCs'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% ILC1_cycling$barcode, 'Cycling group 1 ILCs'))
rownames(bcs) <- bcs$barcode
bcs$barcode <- NULL
ilT <- AddMetaData(ilT, metadata = bcs) # add new annotations to meta data slot
```

## Visualize cell annotations on t-SNE/UMAP

Rearrange the data:

```{r, warning = FALSE, error = FALSE}
Idents(ilT) <- ilT$cellID
levels(ilT) <- c('Naive CD4/CD8 ab T cells', 'Follicular CD4 ab T cells', 'Activated CD4 ab T cells', 
                'Cycling CD4 ab T cells', 'Activated CD8 ab T cells', 'Cytotoxic CD8 ab T cells',
                'Cycling CD8 ab T cells', 'Activated gd T cells', 'Cytotoxic gd T cells', 
                'Cycling gd T cells', 'SELLhi gd T cells', 'CD2neg GD T cells', 'Activated group 1 ILCs',
                'Cytotoxic group 1 ILCs', 'Cycling group 1 ILCs', 'Group 3 ILCs')
ilT$cellID <- Idents(ilT)
```
                
Plot on t-SNE:

```{r, warning = FALSE, error = FALSE}
cols <- c('lightcoral', 'pink', 'violetred1', 'violetred4', 'yellow', 'goldenrod1', 
          'yellow4', 'darkorange', 'darkorange3', 'chocolate4', 'tan', 'sandybrown', 
          'red', 'indianred1', 'darkred', 'black')
DimPlot(ilT, 
        reduction = 'tsne',
        group.by = 'cellID',
        cols = cols)
```

Plot on UMAP:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        reduction = 'umap',
        group.by = 'cellID',
        cols = cols)
```

Plot again, one cell type at a time

Naive ab T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = naiveT, 
        cols.highlight = "lightcoral", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cycling CD4 ab T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = CyclingCD4T$barcode, 
        cols.highlight = "violetred4", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Follicular CD4 ab T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = FollicularCD4T$barcode, 
        cols.highlight = "pink", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Activated CD4 ab T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = ActivatedCD4T$barcode, 
        cols.highlight = "violetred1", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Activated CD8 ab T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = CD8_activated$barcode, 
        cols.highlight = "yellow", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Activated gd T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = gd_activated$barcode, 
        cols.highlight = "darkorange", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

CD2neg gd T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = gd_CD2neg$barcode, 
        cols.highlight = "sandybrown", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cycling CD8 ab T cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = CD8_cycling$barcode, 
        cols.highlight = "yellow4", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cycling gd cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = gd_cycling$barcode, 
        cols.highlight = "chocolate4", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cytotoxic CD8 ab cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = CD8_cytotoxic$barcode, 
        cols.highlight = "goldenrod1", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cytotoxic gd cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = gd_cytotoxic$barcode, 
        cols.highlight = "darkorange3", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

SELLhi gd cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = gd_SELLpos$barcode, 
        cols.highlight = "tan", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Activated group 1 ILCs:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = ILC1$barcode, 
        cols.highlight = "red", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cytotoxic group 1 ILCs:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = ILC1_cytotoxic$barcode, 
        cols.highlight = "indianred1", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cycling group 1 ILCs:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = ILC1_cycling$barcode, 
        cols.highlight = "darkred", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Group 3 ILCs:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilT, 
        cells.highlight = ILC3$barcode, 
        cols.highlight = "black", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

### Save data

Lastly, let's re-save our Seurat object with the new annotations stored in the cellID meta data slot:

```{r, warning = FALSE, error = FALSE, message = FALSE}
Idents(ilT) <- ilT$cellID
saveRDS(ilT, '/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_TILConly.rds')
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References
  
Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).

Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical
  Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/.

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

Hadley Wickham and Jennifer Bryan (2019). readxl: Read Excel Files. R package version 1.3.1.
  https://CRAN.R-project.org/package=readxl
  
Hadley Wickham and Dana Seidel (2020). scales: Scale Functions for Visualization. R package version 1.1.1.
  https://CRAN.R-project.org/package=scales
  
Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data Manipulation.
  R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
