---
title: "Characterization of Ileal Cells: Overall Differential Gene Expression (DGE) Analysis"
author: "Jayne Wiarda"
date: "01June2021"
output: 
  github_document: 
    keep_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created our final annotations for our data and need to re-do overall DGE analysis with these final annotations.

## Overall DGE analysis of cell types

Our criteria for a differentially expressed gene are as follows:

* absolute logFC value > 0.25 (|avg_logFC| > 0.25)
* gene expressed in at least 10% of cells within one of the two populations being compared
  - Population 1 = cell type of interest
  - Population 2 = all other cells in dataset
* adjusted p-value < 0.05

We use the default (Wilcoxon) method to determine DGE and report logFC as the natural log.

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019). To run some of these functions more quickly, we also utilize parallelization with the software package, future (Bengtsson 2020). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

Other required packages: dplyr (Wickham et al. 2020), ggplot2 (Wickham 2016), writexl (Ooms 2020), scales (Wickham & Seidel 2020)

```{r, message = FALSE}
library(Seurat)  
library(writexl)        
library(scales)
library(ggplot2)
library(dplyr)  
library(future)
```

### Import relevant data 

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
il <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/IleumAtlasAll.rds") 
DefaultAssay(il) <- "RNA"
Idents(il) <- il$phylo # place clusters in phylo order determined by hierarchical clustering analysis
```

### Allow parallelization 

```{r, warning = FALSE, error = FALSE}
plan("multiprocess", workers = 7) # change plan to allow parallelization with 7 workers
```

### Perform overall DGE analysis:

```{r, warning = FALSE, error = FALSE, message = FALSE}
ilDE <- FindAllMarkers(il, 
                       only.pos = FALSE, 
                       logfc.threshold = 0.25, # minimum logFC of 0.25
                       min.pct = 0.1, # expressed in at least 10% of all cells in the cluster or entire dataset
                       assay = "RNA") 
ilDE <- subset(ilDE, p_val_adj < 0.05) # make sure the adjusted p-values are still < 0.05 since some genes in DE list have p_val_adj > 0.05
```

We also add in additional gene symbol and Ensembl ID information:

```{r, warning = FALSE, error = FALSE}
features <- read.table("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/QC/GeneInfo.txt", 
                       header = TRUE) # read in the gene information for the filtered gene list from QC analysis
features <- features[c("Symbol", "EnsemblID", "Name")] # subset only the columns of gene symbols, Ensembl IDs, and the names used for analysis
ilDE <- merge(ilDE,
              features, 
              by.x = "gene", 
              by.y = "Name") # merge the DE gene lists with the additional gene information
ilDE <- ilDE[order(ilDE$cluster, ilDE$p_val_adj),] # reorder by lowest to highest p-value within each cluster
```

### Save our differentially expressed gene lists

```{r, warning = FALSE, error = FALSE, eval = TRUE}
write.table(ilDE, 
            file = "/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/DE/AllCells_FINALannot_OverallDE.txt") 
write_xlsx(x = ilDE, 
           path = "/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/DE/AllCells_FINALannot_OverallDE.xlsx",
           col_names = TRUE)
```

### Make a heatmap of the top DE genes

```{r, warning = FALSE, error = FALSE, fig.height = 30, fig.width = 20}
ilDE <- subset(ilDE, avg_logFC > 0) # only take genes enriched in the clusters 
Idents(il) <- il$phylo
topgenes <- ilDE %>% group_by(cluster) %>% top_n(10, avg_logFC) # only plot top 10 genes per cluster, as determined by highest average logFC values
DoHeatmap(subset(il, downsample = 100), # take only 100 cells per cluster for plotting
          features = as.character(topgenes$gene), 
          assay = "RNA", 
          disp.min = -1.5, 
          disp.max = 3) +
  scale_fill_gradientn(colors = c("darkturquoise", "grey90", "indianred1", "red"))
```

As an FYI, DoHeatmap won't run on datasets with >30,000 cells, so must downsample large datasets for heatmap visualization

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References

H. Bengtsson, A Unifying Framework for Parallel and Distributed Processing in R using Futures, arXiv:2008.00553, 2020

Jeroen Ooms (2020). writexl: Export Data Frames to Excel 'xlsx' Format. R package version 1.3.1. https://CRAN.R-project.org/package=writexl
  
Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
 
Hadley Wickham and Dana Seidel (2020). scales: Scale Functions for Visualization. R package version 1.1.1. https://CRAN.R-project.org/package=scales

Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data Manipulation. R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
  
