---
title: "Adding cell annotations to B dataset"
author: "Jayne Wiarda"
date: "20June2021"
output: 
  github_document: 
    keep_html: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created our final annotations for our B data and need to add these and visualize.

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019), ggplot2 (Wickham 2016), scales (Wickham & Seidel 2020), dplyr (Wickham et al. 2020), readxl (Wickham & Bryan 2019), reshape2 (Wickham 2007). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

```{r, message = FALSE}
library(Seurat)
library(ggplot2)
library(scales)
library(dplyr)
library(readxl)
library(reshape2)
```

### Import relevant data

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
ilB <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_Bonly.rds") 
DefaultAssay(ilB) <- "RNA" # set our default assay to RNA for subsequent work
Idents(ilB) <- ilB$seurat_clusters # change our default IDs to our cluster assignments
```

Read in cell annotation files:

```{r, warning = FALSE, error = FALSE}
Bann <- read.table('/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Annotations/B.txt')
```

## Add cell annotations to the dataset

```{r, warning = FALSE, error = FALSE}
bcs <- as.data.frame(colnames(ilB))
colnames(bcs) <- 'barcode'
bcs$cellID <- rep('_', nrow(bcs))
ASC <- subset(Bann, cellID == 'Antibody-secreting cells')
B_cycling <- subset(Bann, cellID == 'Cycling B cells')
B_resting <- subset(Bann, cellID == 'Resting B cells')
B_transitioning <- subset(Bann, cellID == 'Transitioning B cells')
B_nonrestingnoncycling <- subset(Bann, cellID == 'Activated B cells')

bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% ASC$barcode, 'Antibody-secreting cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% B_cycling$barcode, 'Cycling B cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% B_resting$barcode, 'Resting B cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% B_transitioning$barcode, 'Transitioning B cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% B_nonrestingnoncycling$barcode, 'Activated B cells'))
rownames(bcs) <- bcs$barcode
bcs$barcode <- NULL
ilB <- AddMetaData(ilB, metadata = bcs) # add new annotations to meta data slot
```

## Visualize cell annotations on t-SNE/UMAP

Rearrange the data:

```{r, warning = FALSE, error = FALSE}
Idents(ilB) <- ilB$cellID
levels(ilB) <- c('Antibody-secreting cells', 'Transitioning B cells', 'Resting B cells', 
                'Cycling B cells', 'Activated B cells')
ilB$cellID <- Idents(ilB)
```
                
Plot on t-SNE:

```{r, warning = FALSE, error = FALSE}
cols <- c('darkslateblue', 'mediumorchid', 
          'plum4', 'mediumpurple1', 'magenta4')
DimPlot(ilB, 
        reduction = 'tsne',
        group.by = 'cellID',
        cols = cols)
```

Plot on UMAP:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilB, 
        reduction = 'umap',
        group.by = 'cellID',
        cols = cols)
```

Plot again, one cell type at a time

ASCs:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilB, 
        cells.highlight = ASC$barcode, 
        cols.highlight = "darkslateblue", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Transitioning B cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilB, 
        cells.highlight = B_transitioning$barcode, 
        cols.highlight = "mediumorchid", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Resting B cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilB, 
        cells.highlight = B_resting$barcode, 
        cols.highlight = "plum4", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Cycling B cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilB, 
        cells.highlight = B_cycling$barcode, 
        cols.highlight = "mediumpurple1", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Activated B cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilB, 
        cells.highlight = B_nonrestingnoncycling$barcode, 
        cols.highlight = "magenta4", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

### Save data

Lastly, let's re-save our Seurat object with the new annotations stored in the cellID meta data slot:

```{r, warning = FALSE, error = FALSE, message = FALSE}
Idents(ilB) <- ilB$cellID
saveRDS(ilB, '/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_Bonly.rds')
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References
  
Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).

Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical
  Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/.

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

Hadley Wickham and Jennifer Bryan (2019). readxl: Read Excel Files. R package version 1.3.1.
  https://CRAN.R-project.org/package=readxl
  
Hadley Wickham and Dana Seidel (2020). scales: Scale Functions for Visualization. R package version 1.1.1.
  https://CRAN.R-project.org/package=scales
  
Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data Manipulation.
  R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
