---
title: "Adding cell annotations to Myeloid dataset"
author: "Jayne Wiarda"
date: "20June2021"
output: 
  github_document: 
    keep_html: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

We have now created our final annotations for our myeloid data and need to add these and visualize.

### Load required software packages

The following analyses are conducted using base R packages or the scRNA-seq analysis package, Seurat (Stuart and Butler et al. 2019), ggplot2 (Wickham 2016), scales (Wickham & Seidel 2020), dplyr (Wickham et al. 2020), readxl (Wickham & Bryan 2019), reshape2 (Wickham 2007). Refer to sessionInfo() at the bottom of the page for the R and package versions used.

```{r, message = FALSE}
library(Seurat)
library(ggplot2)
library(scales)
library(dplyr)
library(readxl)
library(reshape2)
```

### Import relevant data

Read in the RDS file from previous work:

```{r, warning = FALSE, error = FALSE}
ilM <- readRDS("/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_MyeloidOnly.rds") 
DefaultAssay(ilM) <- "RNA" # set our default assay to RNA for subsequent work
Idents(ilM) <- ilM$seurat_clusters # change our default IDs to our cluster assignments
```

## Add cell annotations to the dataset

```{r, warning = FALSE, error = FALSE}
bcs <- as.data.frame(colnames(ilM))
colnames(bcs) <- 'barcode'
bcs$cellID <- rep('_', nrow(bcs))
df <- as.data.frame(ilM$seurat_clusters)
# Cluster-based cell annotations not provided in the annotation .txt files:
mast <- rownames(subset(df, ilM$seurat_clusters == '52'))
DC <- rownames(subset(df, ilM$seurat_clusters == '42'))
mac <- rownames(subset(df, ilM$seurat_clusters == '49'))

bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% mast, 'Mast cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% DC, 'Dendritic cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% mac, 'Macrophages'))
rownames(bcs) <- bcs$barcode
bcs$barcode <- NULL
ilM <- AddMetaData(ilM, metadata = bcs) # add new annotations to meta data slot
```

## Visualize cell annotations on t-SNE/UMAP

Rearrange the data:

```{r, warning = FALSE, error = FALSE}
Idents(ilM) <- ilM$cellID
levels(ilM) <- c('Dendritic cells', 'Macrophages', 'Mast cells')
ilM$cellID <- Idents(ilM)
```
                
Plot on t-SNE:

```{r, warning = FALSE, error = FALSE}
cols <- c('cornflowerblue', 'cyan4', 'blue')
DimPlot(ilM, 
        reduction = 'tsne',
        group.by = 'cellID',
        cols = cols)
```

Plot on UMAP:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilM, 
        reduction = 'umap',
        group.by = 'cellID',
        cols = cols)
```

Plot again, one cell type at a time

DCs:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilM, 
        cells.highlight = DC, 
        cols.highlight = "cornflowerblue", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Macrophages:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilM, 
        cells.highlight = mac, 
        cols.highlight = "cyan4", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

Mast cells:

```{r, warning = FALSE, error = FALSE}
DimPlot(ilM, 
        cells.highlight = mast, 
        cols.highlight = "blue", 
        sizes.highlight = 1.5,
        cols = "grey85", 
        order = TRUE, 
        reduction = 'tsne') & NoAxes() & NoLegend()
```

### Save data

Lastly, let's re-save our Seurat object with the new annotations stored in the cellID meta data slot:

```{r, warning = FALSE, error = FALSE, message = FALSE}
Idents(ilM) <- ilM$cellID
saveRDS(ilM, '/home/Jayne.Wiarda/scRNAseqIleumAtlas/Ileum/Seurat/Ileum_MyeloidOnly.rds')
```

### View session information

```{r, warning = FALSE, error = FALSE}
sessionInfo()
```

### References
  
Stuart and Butler et al. Comprehensive Integration of Single-Cell Data. Cell (2019).

Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical
  Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/.

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

Hadley Wickham and Jennifer Bryan (2019). readxl: Read Excel Files. R package version 1.3.1.
  https://CRAN.R-project.org/package=readxl
  
Hadley Wickham and Dana Seidel (2020). scales: Scale Functions for Visualization. R package version 1.1.1.
  https://CRAN.R-project.org/package=scales
  
Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data Manipulation.
  R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
